---
title: "consulting project"
date: "2022-11-10"
output: html_document
---

```{r data_setup, message = FALSE, warning = FALSE, echo = FALSE}
#Model building

library(readxl)
library(tidyverse)
library(tree)
library(gam)
library(ggplot2)
library(glmnet)
library(fastDummies)
library(ggcorrplot)
library(gtsummary)

#### data setup
MUA_data = read_excel("TKAMUA5_IDSET_Nov02.xlsx", sheet = "Final", range = cell_rows(3:668), col_names=TRUE)

#Changing MUA/C_MUA in to binary 1 or 0
MUA_data1 = mutate(MUA_data,
  MUA_bi= ifelse(MUA == "Yes", 1, 0),
  C_MUA_bi = ifelse(C_MUA == "Yes", 1, 0),
  MUA_count_T = ifelse(MUA_type == "Both", mua_count +1, mua_count),
  age_diff = age_C_TKA - age,
  los_total = los+los_C_TKA,
  op_time_total = op_time+ op_time_C_TKA,
  date_diff =as.Date(`Date of contralateral TKA`) - as.Date(surgery_date) )

MUA_data1$date_diff <- ifelse(MUA_data1$date_diff <0, -MUA_data1$date_diff,MUA_data1$date_diff)

MUA_data1 = MUA_data1[1:665,]

MUA_data1$C_MUA<- as.factor(MUA_data1$C_MUA)
MUA_data1$MUA<- as.factor(MUA_data1$MUA)

MUA_data1$BTKA <- ifelse(MUA_data1$date_diff==0, "Yes","No")   

# age
MUA_data1$Fac_age_C_TKA <- ifelse(MUA_data1$age_C_TKA<50,"less50",
ifelse(MUA_data1$age_C_TKA<60, "50s",
ifelse(MUA_data1$age_C_TKA<70,"60s","over70s")))

MUA_data1$Fac_age_TKA <- ifelse(MUA_data1$age<50,"less50",
ifelse(MUA_data1$age<60, "50s",
ifelse(MUA_data1$age<70, "60s","over70s")))

# race 
MUA_data1$redu_race <- ifelse(MUA_data1$race %in% c("Hispanic","Multiracial","Other","Preference not indicated","American Indian","Asian"), "Other",MUA_data1$race)

#ethnicity
MUA_data1$ethnicity <- ifelse(MUA_data1$ethnicity %in% c("Prefers not to answer"), "Non-Hispanic Origin",MUA_data1$ethnicity)

#Insurance
names(MUA_data1)[13] <- c("Insurance")
MUA_data1$Insurance <- 
  ifelse(MUA_data1$Insurance %in% c("Blue Cross Commercial","Commercial LUHS","Insurance","Worker's Comp","Worker's Comp LUHS"), "Private",
    ifelse(MUA_data1$Insurance %in% c("Medicare","Medicare LUHS","Managed Medicare"),"Medicare",
      ifelse(MUA_data1$Insurance %in% c("Managed Medicaid","Medicaid","MMAI"),"Medicaid","Uninsured")))

names(MUA_data1)[77] <- c("Insurance_C_TKA")
MUA_data1$Insurance_C_TKA <- 
  ifelse(MUA_data1$Insurance_C_TKA %in% c("Blue Cross Commercial","Commercial LUHS","Insurance"), "Private",
         ifelse(MUA_data1$Insurance_C_TKA %in% c("Medicare","Medicare LUHS","Managed Medicare"),"Medicare",
                ifelse(MUA_data1$Insurance_C_TKA %in% c("Worker's Comp","Worker's Comp LUHS"), "Work_Comp",
                       ifelse(MUA_data1$Insurance_C_TKA %in% c("Managed Medicaid","Medicaid","MMAI"),"Medicaid","Uninsured"))))


### tobacco
MUA_data1$tobacco_C_TKA<- ifelse(is.na(MUA_data1$tobacco_C_TKA)==TRUE, MUA_data1$tobacco,MUA_data1$tobacco_C_TKA) 
MUA_data1$redu_tobacco <- ifelse(MUA_data1$tobacco %in% c("Passive","Yes"), "Yes","No")
MUA_data1$redu_tobacco_C_TKA <- ifelse(MUA_data1$tobacco_C_TKA %in% c("Passive","Yes"), "Yes","No")


### ASA

# missing value 
MUA_data1$ASA<- ifelse(is.na(MUA_data1$ASA)==TRUE, MUA_data1$ASA_C_TKA,MUA_data1$ASA) 

MUA_data1$redu_ASA_C_TKA <- ifelse(MUA_data1$ASA_C_TKA %in% c(2,3),"2/3","else")
MUA_data1$redu_ASA <- ifelse(MUA_data1$ASA %in% c(2,3),"2/3","else")


###los
MUA_data1$los_C_TKA<- ifelse(is.na(MUA_data1$los_C_TKA)==TRUE, MUA_data1$los,MUA_data1$los_C_TKA) 



```

#### Q1: MUA is a risk factor of C_MUA? - Yes

```{r message = FALSE, warning = FALSE}
#Pearson's Chi-squared test
chisq.test(MUA_data1$MUA, MUA_data1$C_MUA)
```

##### \* Mosic plot

```{r echo = FALSE}
Q1_data <- MUA_data1[, 108:109]
colnames(Q1_data)<- c("MUA", "C_MUA" )
Q1_table <- table(Q1_data)
rownames(Q1_table) <- c("No MUA", "MUA")
colnames(Q1_table) <- c("No C_MUA", "C_MUA")

#mosicplot
mosaicplot(Q1_table, main = "MUA plot", color = TRUE)
```

##### \* Fisher's exact test for count data

```{r echo = FALSE, warning = FALSE, message = FALSE}
# fisher's exact test for count data 
test<-fisher.test(Q1_table)

# combine plot and statistical test with ggbarstats
library(ggstatsplot)
ggbarstats(
  Q1_data, MUA, C_MUA,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
)

```

##### \* Logistic regression model

```{r}

#Model with  MUA predicting C_mua
mua_glm = glm(C_MUA ~ MUA, data= MUA_data1, family = "binomial")
summary(mua_glm)

```

-   logit(odds) = -4.5061 + 2.5602\*MUA_bi (0/1)
-   if the patient got MUA, the odds of getting C_MUA is exp(-4.5061 + 2.5602) = 0.1428586
-   if the patient didn't get MUA, the odds of getting C_MUA is exp(-4.5061) = 0.01104144

##### \* Odds Ratio & Correlation

```{r}
#Odds ratio
exp(mua_glm$coefficients[-1])  
```

Odds of receiving MUA on contralateral knee is almost 13x times larger than not

```{r}
cor(MUA_data1$C_MUA_bi,MUA_data1$MUA_bi)
```

------------------------------------------------------------------------

#### Q2: is the \# between 2 TKAs a risk factor? - No

##### \* boxplot of \# days between 2 TKAs

```{r echo = FALSE}

MUA_data1_1 <- MUA_data1 %>%  select(date_diff,C_MUA)
MUA_data1_1 %>% select(date_diff,C_MUA) %>% 
  tbl_summary(by = C_MUA) %>% #, statistic = c(age)~"{mean}({min},{max})")  %>% 
  add_n() %>% modify_header(label ~ "**Variable**") %>%
  bold_labels()

ggplot(data = MUA_data1, aes(x = date_diff, y=C_MUA)) +
  geom_boxplot(aes(fill=C_MUA))+
  #geom_point(aes(group=C_MUA), position = position_dodge(width = 0.75)) +
  coord_flip()+
  labs(x = "# days between 2 TKAs",y = "C_MUA including Bilateral TKAs", fill=NULL)+ theme(legend.position="none")




MUA_data1_1 <- MUA_data1 %>%
  filter(BTKA == "No") %>% 
  select(date_diff,C_MUA)
MUA_data1_1 %>% select(date_diff,C_MUA) %>% 
  tbl_summary(by = C_MUA) %>% #, statistic = c(age)~"{mean}({min},{max})")  %>% 
  add_n() %>% modify_header(label ~ "**Variable**") %>%
  bold_labels()

MUA_data1 %>% filter(BTKA == "No") %>% 
  ggplot(aes(x = date_diff, y=C_MUA)) +        
  geom_boxplot(aes(fill=C_MUA))+
  #geom_point(aes(group=C_MUA), position = position_dodge(width = 0.75)) +
  coord_flip()+
  labs(x = "# days between 2 TKAs",y = "C_MUA excluding Bilateral TKAs", fill=NULL)+ theme(legend.position="none")


```

```{r}
#Model with  MUA predicting C_mua with time

cor(MUA_data1$C_MUA_bi,MUA_data1$date_diff)


mua_glm1 = glm(C_MUA ~ date_diff, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~ date_diff, data= MUA_data1[MUA_data1$BTKA=="No",], family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~ MUA +date_diff, data= MUA_data1, family = "binomial")
summary(mua_glm1)
mua_glm1 = glm(C_MUA ~ (MUA +date_diff)^2, data= MUA_data1, family = "binomial")
summary(mua_glm1)

```

the correlation between days between 2TKAs and C_MUA is close to 0. In logistic regression model, the days between 2 TKAs is insignificant

let's see patients who got the 2nd TKA within 645 days

```{r}
# let's see patients who got the 2nd TKA within 645 days
summary(as.vector(MUA_data1$date_diff))   #3rd Q : 645

cor(MUA_data1[MUA_data1$date_diff<645, ]$C_MUA_bi,MUA_data1[MUA_data1$date_diff<645, ]$date_diff)

mua_glm1 = glm(C_MUA ~ date_diff, data= MUA_data1[MUA_data1$date_diff<645&&MUA_data1$BTKA=="No", ], family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~ MUA+date_diff+ MUA*date_diff, data= MUA_data1[MUA_data1$date_diff<645, ], family = "binomial")
summary(mua_glm1)
# getting obvious that date_diff is insignificant 
```

Even though the correlation between days between 2TKAs and C_MUA changed from negative to positive, it is still close to 0. When we looked into the days within 3rd quartiles, it is getting obvious that date_diff is insignificant

-   If the patient got Bilateral TKA, is the risk of getting MUA significantly higher? : No.

```{r}
# If the patient got Bilateral TKA, is the risk of getting MUA significantly higher? No. 
MUA_data1$BTKA <- ifelse(MUA_data1$date_diff==0, "Yes","No")   
mua_glm1 = glm(as.factor(Group2) ~ as.factor(BTKA), data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~ as.factor(BTKA), data= MUA_data1, family = "binomial")
summary(mua_glm1)
```

-   want to mention about below in the presentation?

```{r echo=FALSE, warning = FALSE, message = FALSE}
# 1yr/2yr/more than 2yr 
MUA_data1$yeardiff <- ifelse(MUA_data1$date_diff<366, "1yr",
                             ifelse(MUA_data1$date_diff<365*2,"2yr","more than 2yrs"))  
#MUA_data1$yeardiff<- ordered(as.factor(MUA_data1$yeardiff), levels = c("1yr", "2yr", "more than 2yrs"))
mua_glm1 = glm(C_MUA ~ yeardiff, data= MUA_data1, family = "binomial")
summary(mua_glm1)


# decision tree 
MUA_tree <- tree(C_MUA ~ MUA +date_diff,data= MUA_data1)
summary(MUA_tree)  # too many nodes and there is no way to prune. 

MUA_tree1 <- tree(C_MUA ~ MUA +date_diff,data= MUA_data1[MUA_data1$date_diff<645, ])
summary(MUA_tree1)

MUA_tree2 <- tree(C_MUA ~ MUA +BTKA,data= MUA_data1)
summary(MUA_tree2)  # BTKA is not used in the decision tree. 

```

------------------------------------------------------------------------

#### Q3 Demographics and comorbidities factors?

##### (1) sex : No

a.  EDA

```{r sexEDA, echo = FALSE}

  ggplot(data = MUA_data1) +
    geom_bar(mapping = aes(x = C_MUA, fill=sex), position = position_dodge(preserve = "single"))+
    labs(x = "# contraletral MUA patients in each sex") 

# proportion of MUA in each gender 
MUA_data1 %>% group_by(sex) %>%  count(C_MUA) %>% 
  mutate(pct = prop.table(n)) %>% 
ggplot(aes(x = C_MUA, y = pct, fill = C_MUA, label = scales::percent(pct))) +
  geom_col( position = "dodge")+
  facet_wrap(~sex) +
  geom_text(position = position_dodge(width = .9),    # move to center of bars
  #  vjust = -0.5,    # nudge above top of bar
    size = 5)+ 
  labs(x = "% of contraletral MUA patients in each sex", y=NULL)+
  theme(legend.position="none")
```

b. chi-squared test 
```{r}
names(MUA_data1)
sex <- MUA_data1[,c(1:4,6)]
sex_table <- table(sex[,4:5])

# fisher's exact test for count data 
test <-chisq.test(sex_table)
fisher.test(sex_table)
fisher.test(table(MUA_data1$C_MUA, MUA_data1$sex))

# combine plot and statistical test with ggbarstats
library(ggplot2)
library(ggstatsplot)
ggbarstats(
  MUA_data1,C_MUA,sex,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(fisher.test(sex_table)$p.value, 3))
  )
) + scale_fill_manual(values=c("#69b3a2", "#404080"))

```



b.  fitting model

```{r sexModel}
##### add sex variable  - insignificant #######
mua_glm1 = glm(C_MUA ~ sex, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 <- glm(as.factor(Group2) ~ sex, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~ MUA +sex, data= MUA_data1, family = "binomial")
summary(mua_glm1)
```

The sex variable is insignificant in these models.

##### (2) Age

a.  EDA

```{r ageEDA, echo = FALSE}
# histogram
ggplot(data = MUA_data1, aes(x = age_C_TKA, fill=C_MUA)) +
  geom_histogram(binwidth = 4,color="#e9ecef")+
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  labs(y = NULL, fill=NULL)+ theme(legend.position="none")

# boxplot 
ggplot(data = MUA_data1, aes(x = age_C_TKA, y = C_MUA)) +
  geom_boxplot(aes(fill=C_MUA))+
 # geom_point(aes(group=C_MUA), position = position_dodge(preserve = "single")) +
  coord_flip()+
  labs(y = NULL, fill=NULL)+ theme(legend.position="none")

# proportion of C_MUA when the patients are in 50s or 60s 8/391 : 0.02046036
MUA_data1 %>% filter(age_C_TKA >50 & age_C_TKA <70) %>%  count(C_MUA) %>% # No:391/Yes: 8
  mutate(pct = prop.table(n)) %>% 
  ggplot(aes(x = C_MUA, y = pct, fill = C_MUA, label = scales::percent(pct))) +
  geom_col( position = "dodge")+
  #facet_wrap(~ MUA) +
  geom_text(position = position_dodge(width = .9),    # move to center of bars
    #  vjust = -0.5,    # nudge above top of bar
    size = 5)+ 
  labs(x = "Contraletral MUA when in 50s or 60s", y=NULL)+
  theme(legend.position="none")

# proportion of C_MUA when the patients are in 40s or over 70s  2/241 : 0.008298755
MUA_data1 %>% filter(age_C_TKA <50 | age_C_TKA >70) %>%  count(C_MUA) %>%  # No:241/Yes: 2
  mutate(pct = prop.table(n)) %>% 
  ggplot(aes(x = C_MUA, y = pct, fill = C_MUA, label = scales::percent(pct))) +
  geom_col( position = "dodge")+
  #facet_wrap(~ MUA) +
  geom_text(position = position_dodge(width = .9),    # move to center of bars
    #  vjust = -0.5,    # nudge above top of bar
    size = 5)+ 
  labs(x = "Contraletral MUA when in younger than 50s or older than 60s", y=NULL)+
  theme(legend.position="none")


MUA_data1 %>%  group_by(Fac_age_C_TKA) %>% count(C_MUA) %>%  # No:241/Yes: 2
  mutate(pct = prop.table(n)) %>% 
  ggplot(aes(x = C_MUA, y = pct, fill = C_MUA, label = scales::percent(pct))) +
  geom_col( position = "dodge")+
  facet_wrap(~ Fac_age_C_TKA) +
  geom_text(position = position_dodge(width = .9),    # move to center of bars
    #  vjust = -0.5,    # nudge above top of bar
    size = 5)+ 
  labs(x = "% of Contraletral MUA patients in each age group", y=NULL)+
  theme(legend.position="none")


```

b.
```{r}
# fisher's exact test for count data 
test <- fisher.test(table(MUA_data1$C_MUA, MUA_data1$Fac_age_C_TKA))

# combine plot and statistical test with ggbarstats
library(ggstatsplot)
ggbarstats(
  MUA_data1,C_MUA,Fac_age_C_TKA,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
) + scale_fill_manual(values = c( "darkturquoise","coral"))

```


c.  regression model

```{=html}
<!-- -->
```
(i) with numerical age variable

```{r}
mua_glm1 = glm(C_MUA ~ age_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~ MUA +age_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)  #exp(-0.04859) = 0.9525716 : as age increase 1, the odds of getting C_MUA decrease 5%..? Not true
# numeric age variable seems insignificant and not to make sense. EDA shows that when the patients are in 50s, 60s, the odds of getting MUA seems higher 
```

(ii) with factorized age variable

```{r}

MUA_data1$Fac_age_C_TKA <- ifelse(MUA_data1$age_C_TKA<50,"less50",
                                   ifelse(MUA_data1$age_C_TKA<60, "50s",
                                     ifelse(MUA_data1$age_C_TKA<70, "60s","over70s")))

MUA_data1$Fac_age_C_TKA2 <- ifelse(MUA_data1$Fac_age_C_TKA %in% c("50s","60s"),"50/60","younger than 50|older than 60")

MUA_data1 %>% group_by(Fac_age_C_TKA2) %>%  count(Group2)

mua_glm1 = glm(C_MUA ~ Fac_age_C_TKA2, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~ MUA +Fac_age_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~ (MUA +Fac_age_C_TKA2)^2, data= MUA_data1, family = "binomial")
summary(mua_glm1) # interaction term is insignificant 





```

-   when the patients are in 50s or 60s, age is significant. interaction is insignificant
-   keep age in the model.

##### (3) Race

a.  EDA

```{r RaceEDA, echo = FALSE}
ggplot(data = MUA_data1) +
  geom_bar(mapping = aes(fill = redu_race, x=C_MUA),position = position_dodge(preserve = "single"))+
  labs(x = NULL, title = "Race")


# proportion of C_MUA by race 
MUA_data1 %>% group_by(redu_race) %>%  count(C_MUA) %>% 
  mutate(pct = prop.table(n)) %>% 
  ggplot(aes(x = C_MUA, y = pct, fill = C_MUA, label = scales::percent(pct))) +
  geom_col( position = "dodge")+
  facet_wrap(~redu_race) +
  geom_text(position = position_dodge(width = .9),    # move to center of bars
    #  vjust = -0.5,    # nudge above top of bar
    size = 3)+ 
  labs(x = NULL, y=NULL)+
  theme(legend.position="none")
```

b. chi-squared test 
```{r}
# fisher's exact test for count data 
test<- fisher.test(table(MUA_data1$C_MUA, MUA_data1$redu_race))

# combine plot and statistical test with ggbarstats
ggbarstats(
  MUA_data1,C_MUA,redu_race,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
) + scale_fill_manual(values = c( "darkturquoise","coral"))


```



b.  fitting model

```{r raceModel}
#############################  race variable
# the stand.error is way tooooooooo big..
# race black is significant..? 

mua_glm1 <- glm(C_MUA ~ as.factor(redu_race), data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~ MUA+ as.factor(redu_race), data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~ (MUA+ as.factor(redu_race))^2, data= MUA_data1, family = "binomial")
summary(mua_glm1)

#mua_glm1 = glm(C_MUA ~ MUA+ as.factor(redu_race)+Fac_age_C_TKA, data= #MUA_data1, family = "binomial")
#summary(mua_glm1)
```

##### (4). Ethnicity

a.  EDA

```{r EthnicityEDA, echo=FALSE}
ggplot(data = MUA_data1) +
  geom_bar(mapping = aes(x = C_MUA, fill=ethnicity),position = position_dodge(preserve = "single"))+
  labs(x = NULL, title = "Ethnicity" ) 

# proportion of C_MUA in each ethnicity
MUA_data1 %>% group_by(ethnicity) %>%  count(C_MUA) %>% 
  mutate(pct = prop.table(n)) %>% 
  ggplot(aes(x = C_MUA, y = pct, fill = C_MUA, label = scales::percent(pct))) +
  geom_col( position = "dodge")+
  facet_wrap(~ethnicity) +
  geom_text(position = position_dodge(width = .9),    # move to center of bars
    #  vjust = -0.5,    # nudge above top of bar
    size = 5)+ 
  labs(x = "proportion of C_MUA in each ethnicity", y=NULL)+
  theme(legend.position="none")

```

b. chi-squared test 
```{r}
# fisher's exact test for count data 
test<- fisher.test(table(MUA_data1$C_MUA, MUA_data1$ethnicity))

# combine plot and statistical test with ggbarstats
ggbarstats(
  MUA_data1,C_MUA,ethnicity,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
) + scale_fill_manual(values = c( "darkturquoise","coral"))


```



C. regression model

```{r EthnicityModel}
############################  ethnicity variable
mua_glm1 = glm(C_MUA ~  ethnicity, data= MUA_data1, family = "binomial")
summary(mua_glm1)
mua_glm1 = glm(C_MUA ~  MUA+ethnicity, data= MUA_data1, family = "binomial")
summary(mua_glm1)
# the stand.error is tooooooooo big..
```


##### (5) Insurance
a.EDA
```{r InsuranceEDA, echo =FALSE}
# when the 2nd TKA 
ggplot(data = MUA_data1) +
  geom_bar(mapping = aes(x = C_MUA, fill=Insurance_C_TKA),position = position_dodge(preserve = "single"))+
  labs(x = NULL) 

# proportion of MUA in each insurance type
MUA_data1 %>% group_by(Insurance_C_TKA) %>%  count(C_MUA) %>% 
  mutate(pct = prop.table(n)) %>% 
  ggplot(aes(x = C_MUA, y = pct, fill = C_MUA, label = scales::percent(pct))) +
  geom_col( position = "dodge")+
  facet_wrap(~Insurance_C_TKA) +
  geom_text(position = position_dodge(width = .9),    # move to center of bars
    #  vjust = -0.5,    # nudge above top of bar
    size = 3)+ 
  labs(x = NULL, y=NULL)+
  theme(legend.position="none")
```

b. chi-squared test 
```{r}
# fisher's exact test for count data 
test<- fisher.test(table(MUA_data1$C_MUA, MUA_data1$Insurance_C_TKA))

# combine plot and statistical test with ggbarstats
ggbarstats(
  MUA_data1,C_MUA,Insurance_C_TKA,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
) + scale_fill_manual(values = c( "darkturquoise","coral"))


```


 c. Regression model 
```{r InsuranceModel}
mua_glm1 = glm(C_MUA ~  Insurance_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~  MUA+Insurance_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
# decision tree 
MUA_tree <- tree(C_MUA ~ MUA +redu_race+Insurance_C_TKA+Fac_age_C_TKA,data= MUA_data1)
#summary(MUA_tree) # MUA is the only variable used ..
```

##### (6) BMI
a. EDA
```{r BMIEDA, echo=FALSE}

MUA_data1_1 <- MUA_data1 %>%  select(bmi_C_TKA,C_MUA)
MUA_data1_1 %>% select(bmi_C_TKA,C_MUA) %>% 
  tbl_summary(by = C_MUA) %>% #, statistic = c(age)~"{mean}({min},{max})")  %>% 
  add_n() %>% modify_header(label ~ "**Variable**") %>%
  bold_labels()



ggplot(data = MUA_data1, aes(x = bmi_C_TKA, y = C_MUA)) +
  geom_boxplot(aes(fill=C_MUA))+
  #geom_point(aes(group=C_MUA), position = position_dodge(preserve = "single")) +
  coord_flip()+
  labs(y = "Contralateral MUA", fill=NULL)+ theme(legend.position="none")


```

b. Regression model 
```{r BMIModel}
mua_glm1 = glm(C_MUA ~  bmi_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~  MUA+bmi_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)


```


##### (7) tobacco 
a. EDA
```{r tobaccoEDA, echo=FALSE}
  ggplot(data = MUA_data1) +
  geom_bar(mapping = aes(fill = redu_tobacco_C_TKA, x=C_MUA), position = position_dodge(preserve = "single"))+
  labs(x = NULL, fill=NULL)

```


b. Regression Model 
```{r tobaccoModel}
mua_glm1 = glm(C_MUA ~  redu_tobacco_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~  MUA+redu_tobacco_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)

```


##### (8) ASA
a. EDA
```{r ASAEDA, echo = FALSE}
# histogram
ggplot(data = MUA_data1, aes(x = ASA_C_TKA, fill=C_MUA)) +
  geom_histogram(binwidth = 1,color="#e9ecef")+
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  labs(y = NULL, fill=NULL)+ theme(legend.position="none")

# proportion of MUA in each ASA rate
MUA_data1 %>% group_by(ASA_C_TKA) %>%  count(C_MUA) %>% 
  mutate(pct = prop.table(n)) %>% 
  ggplot(aes(x = C_MUA, y = pct, fill = C_MUA, label = scales::percent(pct))) +
  geom_col( position = "dodge")+
  facet_wrap(~ASA_C_TKA) +
  geom_text(position = position_dodge(width = .9),    # move to center of bars
    #  vjust = -0.5,    # nudge above top of bar
    size = 3)+ 
  labs(x = "proportion of MUA in each ASA", y=NULL)+
  theme(legend.position="none")
```

b. fisher
```{r}
# fisher's exact test for count data 
test<- fisher.test(table(MUA_data1$C_MUA, MUA_data1$ASA_C_TKA))

# combine plot and statistical test with ggbarstats
ggbarstats(
  MUA_data1,C_MUA,ASA_C_TKA,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
) + scale_fill_manual(values = c( "darkturquoise","coral"))
```

b. Regression model
```{r ASAmodel}
mua_glm1 = glm(C_MUA ~ as.factor(ASA_C_TKA), data= MUA_data1, family = "binomial")
summary(mua_glm1)
# standard error is too big so reduced the factors into binary

MUA_data1$redu_ASA_C_TKA <- ifelse(MUA_data1$ASA_C_TKA %in% c(2,3),"2/3","else")

mua_glm1 = glm(C_MUA ~  as.factor(redu_ASA_C_TKA), data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~  MUA+ as.factor(redu_ASA_C_TKA), data= MUA_data1, family = "binomial")
summary(mua_glm1)


```


#####  (9) los
a. EDA
```{r losEDA, echo=FALSE}

MUA_data1_1 <- MUA_data1 %>%  select(los_C_TKA,C_MUA)
MUA_data1_1 %>% select(los_C_TKA,C_MUA) %>% 
  tbl_summary(by = C_MUA) %>% #, statistic = c(age)~"{mean}({min},{max})")  %>% 
  add_n() %>% modify_header(label ~ "**Variable**") %>%
  bold_labels()


ggplot(data = MUA_data1, aes(x = los_C_TKA, y=C_MUA)) +
  geom_boxplot(aes(fill=C_MUA))+
  #geom_point(aes(group=C_MUA), position = position_dodge(width = 0.75)) +
  coord_flip()+
  labs(y = "C_MUA", fill=NULL)+ theme(legend.position="none")

```


b. regression model 
```{r losModel}
mua_glm1 = glm(C_MUA ~  los_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~  MUA + los_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~  (MUA + los_C_TKA)^2, data= MUA_data1, family = "binomial")
summary(mua_glm1)

```


##### (10) operation time 

a. EDA
```{r optimeEDA, echo = FALSE}

MUA_data1_1 <- MUA_data1 %>%  select(op_time_C_TKA,C_MUA)
MUA_data1_1 %>% select(op_time_C_TKA,C_MUA) %>% 
  tbl_summary(by = C_MUA) %>% #, statistic = c(age)~"{mean}({min},{max})")  %>% 
  add_n() %>% modify_header(label ~ "**Variable**") %>%
  bold_labels()

ggplot(data = MUA_data1, aes(x = op_time_C_TKA, y=C_MUA)) +
  geom_boxplot(aes(fill=C_MUA))+
  #geom_point(aes(group=C_MUA), position = position_dodge(width = 0.75)) +
  coord_flip()+
  labs(x = "2nd TKA Operation Time (mins)",y = NULL, fill=NULL)+ theme(legend.position="none")

```


b. Regression model 
```{r optimeModel}
mua_glm1 = glm(C_MUA ~  op_time_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)

mua_glm1 = glm(C_MUA ~  MUA+op_time_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)

```

##### (11) readmission 

a. EDA
```{r readmissionEDA, echo = FALSE}
MUA_data1$readmit_90d_C_TKA <- as.factor(MUA_data1$readmit_90d_C_TKA)

MUA_data1_1 <- MUA_data1 %>%  select(readmit_90d_C_TKA,C_MUA)
MUA_data1_1 %>% select(readmit_90d_C_TKA,C_MUA) %>% 
  tbl_summary(by = C_MUA) %>% #, statistic = c(age)~"{mean}({min},{max})")  %>% 
  add_n() %>% modify_header(label ~ "**Variable**") %>%
  bold_labels()


  ggplot(data = MUA_data1) +
    geom_bar(mapping = aes(x = C_MUA, fill=readmit_90d_C_TKA), position = position_dodge(preserve = "single"))+
    labs(x = "# contraletral MUA patients")  +
    theme(legend.position = c(0.8, 0.8)) + 
    scale_fill_discrete(labels=c('No readmission', 'readmission'))
  



```
b. fisher
```{r}
# fisher's exact test for count data 
test<- fisher.test(table(MUA_data1$C_MUA, MUA_data1$readmit_90d_C_TKA))

# combine plot and statistical test with ggbarstats
ggbarstats(
  MUA_data1,C_MUA,readmit_90d_C_TKA,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
) + scale_fill_manual(values = c( "darkturquoise","coral"))
```

c. Regression model
```{r Readmissionmodel}
mua_glm1 = glm(C_MUA ~  readmit_90d_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)


mua_glm1 = glm(C_MUA ~  MUA+ readmit_90d_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)
```

##### (13) ER visit within 90 days 


a. EDA
```{r EREDA, echo = FALSE}
MUA_data1$ed_90d_C_TKA <- as.factor(MUA_data1$ed_90d_C_TKA)

MUA_data1_1 <- MUA_data1 %>%  select(ed_90d_C_TKA,C_MUA)
MUA_data1_1 %>% select(ed_90d_C_TKA,C_MUA) %>% 
  tbl_summary(by = C_MUA) %>% #, statistic = c(age)~"{mean}({min},{max})")  %>% 
  add_n() %>% modify_header(label ~ "**Variable**") %>%
  bold_labels()

  ggplot(data = MUA_data1) +
    geom_bar(mapping = aes(x = C_MUA, fill=ed_90d_C_TKA), position = position_dodge(preserve = "single"))+
    labs(x = "# contraletral MUA patients")  +
    theme(legend.position = c(0.8, 0.8)) + 
    scale_fill_discrete(labels=c('No ER', 'ER'))
  

```

b. fisher
```{r}
# fisher's exact test for count data 
test<- fisher.test(table(MUA_data1$C_MUA, MUA_data1$ed_90d_C_TKA))

# combine plot and statistical test with ggbarstats
ggbarstats(
  MUA_data1,C_MUA,ed_90d_C_TKA,
  results.subtitle = FALSE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
) + scale_fill_manual(values = c( "darkturquoise","coral"))
```

C. Regression model
```{r ERmodel, message = FALSE, warning = FALSE}
mua_glm1 = glm(C_MUA ~  ed_90d_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)


mua_glm1 = glm(C_MUA ~  MUA+ ed_90d_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm1)
```




##### (12) comorbidities 
a. EDA
```{r message = FALSE, warning = FALSE}
# summary
Comor_ratio_C_MUA <- apply(MUA_data1[MUA_data1$C_MUA=="Yes",c(82:89,91:101,103)],2,sum)/10
Comor_ratio_NO_C_MUA<- round(apply(MUA_data1[MUA_data1$C_MUA!="Yes",c(82:89,91:101,103)],2,sum)/655,2)
cbind(Comor_ratio_C_MUA,Comor_ratio_NO_C_MUA)

```


b. Lasso

```{r ComorbiditiesLasso, message = FALSE, warning = FALSE}
#Lasso model  - comorbidities 
library(glmnet)
x1matrix = data.matrix(MUA_data1[,c(82:89,91:101,103)]) 
colnames(x1matrix)<- c(str_sub(colnames(x1matrix),end=-7))



C_MUA = as.vector(MUA_data1$C_MUA_bi)

#Finding lambda
k = 5
set.seed(987)
cv.lasso<- cv.glmnet(x1matrix, C_MUA, alpha=1, family = "binomial")
lasso.mod = glmnet(x1matrix, C_MUA, alpha=1,family = "binomial",lambda=cv.lasso$lambda.min)
lasso.coef <- predict(lasso.mod, type = "coefficients", s=cv.lasso$lambda.min)
paste0(round(lasso.coef@x,4),"X",lasso.coef@i, collapse=" + ")

```
Lasso regression model picked blood transfusion and Hemiplegia

* correlation between diagnosed disease and contralateral MUA 
```{r comoCorrelationtable, echo = FALSE, warning = FALSE}
library(ggcorrplot)
ggcorrplot(cor(cbind(x1matrix, C_MUA)),colors = c("#6D9EC1", "white", "#E46726"),type = "lower")
```
The predictor that lasso picked matches with correlation between C_MUA and diseases

#### So final model is..?

```{r finalModel, eval=FALSE, warning=FALSE, include=FALSE}
mua_glm <- glm(C_MUA ~ MUA+Fac_age_C_TKA+MUA_data1$redu_race+los_C_TKA+blood_transfusion_C_TKA+Hemiplegia_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm)


mua_glm <- glm(C_MUA ~ MUA+Fac_age_C_TKA+MUA_data1$redu_race+blood_transfusion_C_TKA+Hemiplegia_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm)

mua_glm <- glm(C_MUA ~ MUA+Fac_age_C_TKA+blood_transfusion_C_TKA+Hemiplegia_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm)

mua_glm <- glm(C_MUA ~ MUA+Fac_age_C_TKA+blood_transfusion_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm)


mua_glm <- glm(C_MUA ~ MUA+blood_transfusion_C_TKA+Hemiplegia_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm)


mua_glm <- glm(C_MUA ~ MUA+blood_transfusion_C_TKA, data= MUA_data1, family = "binomial")
summary(mua_glm)




```

##### (1) LASSO 
###### a. model 
```{r message = FALSE, warning = FALSE, eval = FALSE}
library(glmnet)
names(MUA_data1)
MUA_data2 = MUA_data1[,c(109,108,6,9,75:78, 80:89,91:104,106,116,118)]
MUA_data2$ASA_C_TKA<-as.factor(MUA_data2$ASA_C_TKA)
y = MUA_data2$C_MUA_bi

# Fac_age_C_TKA(116) is included instead of age_C_TKA(74), Diabetes_cc_C_TKA(90) is removed b/c it's duplicated with Diabetes_no_cc_C_TKA
# 108: MUA_bi,  6 : sex, 9: ethnicity, 75: bmi_C_TKA, 76: tobacco_C_TKA,  77: Insurance_C_TKA, 78: los_C_TKA, 80:ASA_C_TKA, 81: op_time_C_TKA, 82~103: comorbidities, 104:readmit_90d_C_TKA, 106:ed_90d_C_TKA,116:Fac_age_C_TKA, 118: redu_race       

d <- as.data.frame(MUA_data2)
options(na.action="na.pass")
m <- model.matrix(C_MUA_bi ~ ., data=d)[,-1]


set.seed(1234)
cv.out = cv.glmnet(m, y, alpha=1,family="binomial")
bestlam = cv.out$lambda.min

lasso.mod = glmnet(m, y, alpha= 1, lambda = bestlam,family="binomial")
coef(lasso.mod)
paste0(round(coef(lasso.mod)@x,4),"X",coef(lasso.mod)@i, collapse=" + ")


##### ROC
#lasso.mod = lasso mod using bestlam from cv.glmnet
#m = original model matrix
predict_fit = predict(lasso.mod, m, type = "response")

library(pROC)
roc_score=roc(predictor=as.vector(predict_fit), response = y ) 

#ROC Plot
plot(roc_score ,main ="ROC curve -Lasso Regression") 
roc_score$auc #AUC score

##### Bootstrapping AUC to find 95% CI
#Finding Bootstrapped AUC confidence intervals
library(fbroc)
boot_roc = boot.roc(as.numeric(predict_fit), as.logical(as.numeric(y)), n.boot = 10000)
plot(boot_roc)

w=perf(boot_roc, "auc") #Measuring performance
w #AUC confidence interval
hist(w$boot.results, main = "Bootstrapped AUC Values", xlab = "Area Under Curve") #Histogram of bootstrapped AUC

```

-4.6632 + 1.3594MUA_bi + 1.2333Insurance_C_TKAWork_Comp + 2.1177blood_transfusion_C_TKA


###### b. leave one out cross validation 
```{r message = FALSE, warning = FALSE, eval=FALSE}
library(glmnet)
#y = MUA_data1$C_MUA_bi
#MUA_data2 = MUA_data1[,c(108, 6, 9,75:78, 80:89,91:104,106,116,118)] # Fac_age_C_TKA(116) is included instead of age_C_TKA(74), Diabetes_cc_C_TKA(90) is removed b/c it's duplicated with Diabetes_no_cc_C_TKA
# 108: MUA_bi,  6 : sex, 9: ethnicity, 75: bmi_C_TKA, 76: tobacco_C_TKA,  77: Insurance_C_TKA, 78: los_C_TKA, 80:ASA_C_TKA, 81: op_time_C_TKA, 82~103: comorbidities, 104:readmit_90d_C_TKA, 106:ed_90d_C_TKA,116:Fac_age_C_TKA, 118: redu_race       
#MUA_data2$ASA_C_TKA<-as.factor(MUA_data2$ASA_C_TKA)

#d <- data.frame(x=MUA_data2, y=y)
#m <- model.matrix(y ~ ., data=d)[,-1]

pred_prob <- c()
for(i in 1:nrow(d)){
y_train = d[-i,]$C_MUA_bi
x_train = m[-i,]

y_test = d[i,]$C_MUA_bi
x_test = m[i,]

set.seed(1234)
cv.out = cv.glmnet(x_train, y_train, alpha =1,family="binomial", nfolds=5)
bestlam = cv.out$lambda.min

lasso.mod = glmnet(x_train, y_train, alpha= 1, lambda = bestlam,family="binomial")
pred_prob[i] <- predict(lasso.mod,s = bestlam, newx = x_test,type="response")
}


pred_C_MUA<-cbind(pred_prob,y)
median(pred_C_MUA[y==0,1])
median(pred_C_MUA[y==1,1])



library(pROC)
roc_score_predictpower=roc(predictor=as.vector(pred_prob), response = y ) 

#ROC Plot
plot(roc_score_predictpower ,main ="ROC curve -Lasso Regression(LOOCV) ") 
roc_score_predictpower$auc #AUC score

```







































